---
title: "Survey Statistics | Project NHAMCS"
author: "Rianne, Piet, Zhenwei, Marco"
date: "Dec 16, 2019"
output:
  html_document:
    code_folding: hide
    theme: cosmo
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---


```{r}
library(tidyverse)
library(survey)
library(sampling)
library(kableExtra)
```


```{r}
data <- read_csv("nhamcsed2010-short.csv") %>%
  rownames_to_column("id")
```


### 1. The variable CPSUM represents geographic primary sampling units (clusters), a variable CSTRATM represents relevant strata (e.g., emergency service areas within emergency departments or outpatient departments), variable PATWT represents an additional patient visit weight variable.

#### a. Explain this multistage sampling design. Do the PATWT weights sum up to N, and what does this mean?

```{r}
sum(data$PATWT)
```

Yes, the weights add up to N.

#### b. Define the svydesign() function with clusters, strata, and (additional) weights.

```{r}
data.design <- svydesign(id = ~CPSUM, strata = ~CSTRATM, weight = ~PATWT, data = data, nest = TRUE)
```

#### c. When considering gender differences in emergency visits, is it more likely for men than women to visit an emergency department, explain?

```{r}
# Population estimates --> 
svytotal(~interaction(SEX), data.design)
```


#### d. Is there a mean difference between males and females in weights (PATWT), what do you conclude from this?

### 2. Estimate the average waiting time in the population under the specified svydesign. 

show the r code line of the specified design. 

#### a. What do you conclude?

```{r}
(result.2a <- svymean(~WAITTIME, data.design, na.rm = T))

prop.table(table(data$WAITTIME))[2]
```
The average waiting time is almost 45 min. with a SE of 1.8512 --> interpretation
Show the proportion of missings --> MCAR? about 6.6. percent, and 2.7 percent not applicable (2.7 have probably too long or felt better) --> 6.6 percent can be imputed with proportion technique about 5 decreasesd show in presentation -_> comparison


#### b. Exclude the cluster and strata information, (redefine the svydesign), how this change your result in 2a, explain?

```{r}
# Defining the design object but without any design specifications
data.design.exc <- svydesign(id = ~id, weight = ~PATWT, data = data) 

(result.2b <- svymean(~WAITTIME, data.design.exc, na.rm = T))
```

The estimates are the same (about 45 min) but the the SE are too small for the option without any design specifications. Explanation needed.

#### c. Draw an SRS sample to estimate the average waiting time. Estimate the average waiting time, and compare the result(s) with the result(s) of 2a and 2b.


 --> sample size stick to 2000 -_> ask students and peter about this
Implement formulas??

```{r}
# With the srswr function we can draw a srs, which gives us a vector of the sample cases
sample.2c.vec.1 <- srswr(n = 1000, N = length(data$id))
sample.2c.vec.2 <- srswr(n = 5000, N = length(data$id))
sample.2c.vec.3 <- srswr(n = 10000, N = length(data$id))
sample.2c.1 <- getdata(data, sample.2c.vec.1)
sample.2c.2 <- getdata(data, sample.2c.vec.2)
sample.2c.3 <- getdata(data, sample.2c.vec.3)

# We can then create another svydesign object and estimate the average waiting time
sample.2c.des.1 <- svydesign(id = ~id, weight = ~PATWT, data = sample.2c.1, nest = TRUE)
sample.2c.des.2 <- svydesign(id = ~id, weight = ~PATWT, data = sample.2c.2, nest = TRUE)
sample.2c.des.3 <- svydesign(id = ~id, weight = ~PATWT, data = sample.2c.3, nest = TRUE)

result.2c.1 <- svymean(~WAITTIME, sample.2c.des.1, na.rm = T)
result.2c.2 <- svymean(~WAITTIME, sample.2c.des.2, na.rm = T)
result.2c.3 <- svymean(~WAITTIME, sample.2c.des.3, na.rm = T)

# Comparing the results 
name.results <- c("2a Estimate (design specified)", "2b Estimate (design not specified)", "2c Estimate (new srs n = 1000)",
                  "2c Estimate (new srs n = 5000)", "2c Estimate (new srs n = 10000)")

# Still need to add the SEs to the table for better comparison
results.2 <- tibble(name.results, estimates = c(result.2a, result.2b, result.2c.1, result.2c.2, result.2c.3),
                    SE = c(map_dbl(list(result.2a, result.2b, result.2c.1, result.2c.2, result.2c.3), 
                                   ~sqrt(attr(., which = "var")))))

```


#### d. Draw a stratified sample to investigate the variability in waiting times over regions (Northeast, Midwest, South, West). Is there variability in waiting times over regions?


```{r}
# In order for drawing a stratified sample we need to arrange the original dataset in ascending strata order (Region)
data.strata.prep <- data %>%
  arrange(REGION)

# With the strata function we can draw a stratified srs. The size vector specifies the strata sample sizes 
# Implemented PPS strata sizes

strata.sizes <- data %>%
  group_by(REGION) %>%
  summarise(prop = (n() / 34936)) %>%
  mutate(size = round(prop * 5000)) # --> this number can change due to sample sizes decided above

sample.2d.vec <- strata(data = data.strata.prep, stratanames = "REGION", size = strata.sizes$size, method = "srswr")
sample.2d <- getdata(data.strata.prep, sample.2c.vec)

sample.2d.des <- svydesign(id = ~id, data = sample.2d, nest = TRUE) # is nest necessary?
svymean(~WAITTIME, sample.2d.des, na.rm = T)

svyby(~WAITTIME, ~REGION, sample.2d.des, svymean, deff = T)
```

imagine we want to do this study again but only for estimation of waiting time 

the variance between the strata is higher then the variance within in strata --> strata are more similar within in then between strata --> oversample regions with higher variance for the next question

### 3. Explain that a ratio estimator can be used to improve the estimation of the average waiting time. 

#### b. Describe potential variables for ratio estimation, and choose one to compute a ratio estimate of the average weighting time. 

we describe multiple ones but we will probably choose the ambulance one

#### c. Give a 95% confidence interval for your preferable estimate of the average waiting time, and explain the result.



### 4. Investigate differences in waiting times among ethnic groups by fitting a linear regression model. Compare estimated results of a stratified sampling design and an unequal probability sampling design.

two models: 
are the individual weights the defined survey weights?
are the stratified weights in comparison to the individual weights related to ethnicity? What do we base the strata on? are these the same as beforehand or are these new strata that we ourselves need to define?
is it just comparing a model with the cluster and strata info like in the beginning and one model without any info


### 5. Assume unequal probability sampling, can you identify the important predictor variables of waiting times of patients visiting an ED? What can be concluded?


We dont know the severity of the patients, probably the most important predictor..


one or two issues. not everyone has to present, slides are nice answering questions 

focus more on the logic of the analysis and have a small (methodologically focused) storyline