---
title: "Survey Statistics | Project NHAMCS"
author: "Rianne, Piet, Zhenwei, Marco"
date: "Dec 16, 2019"
output:
  html_document:
    code_folding: hide
    theme: cosmo
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---


```{r}
library(tidyverse)
library(survey)
library(sampling)
library(kableExtra)
library(svglite)
'%!in%' <- function(x,y)!('%in%'(x,y))

set.seed(3)
```


```{r}
data <- read_csv("nhamcsed2010-short.csv") %>%
  rownames_to_column("id")
```


### 1. The variable CPSUM represents geographic primary sampling units (clusters), a variable CSTRATM represents relevant strata (e.g., emergency service areas within emergency departments or outpatient departments), variable PATWT represents an additional patient visit weight variable.

#### a. Explain this multistage sampling design. Do the PATWT weights sum up to N, and what does this mean?

```{r}
sum(data$PATWT)
```

Yes, the weights add up to N.

#### b. Define the svydesign() function with clusters, strata, and (additional) weights.

```{r}
data.design <- svydesign(id = ~CPSUM, strata = ~CSTRATM, weight = ~PATWT, data = data, nest = TRUE)
```

#### c. When considering gender differences in emergency visits, is it more likely for men than women to visit an emergency department, explain?

```{r}
# Males (2) and Femals (1) proportion
data %>% 
  group_by(SEX) %>% 
  summarise(n = n(), mean(PATWT))

# Population estimates for Males and Females 
svytotal(~interaction(SEX), data.design) 
```


#### d. Is there a mean difference between males and females in weights (PATWT), what do you conclude from this?

### 2. Estimate the average waiting time in the population under the specified svydesign. 

show the r code line of the specified design. 

#### a. What do you conclude?

```{r}
(result.2a <- svymean(~WAITTIME, data.design, na.rm = T))
```

Imputing data by Zhenwei
```{r}
data.imputed <- data %>%
  mutate(WAITTIME.prop = case_when(WAITTIME != -9 & WAITTIME != -7 & LOV != -9 ~ WAITTIME/LOV,
                                 TRUE ~ NA_real_)) %>%
  mutate(WAITTIME.imputed = case_when(WAITTIME != -9 & WAITTIME != -7 ~ WAITTIME,
                                      WAITTIME == -7 ~ -7,
                                      WAITTIME == -9 & LOV != -9 ~ LOV * base::mean(WAITTIME.prop, na.rm= T)))
prop.table(table(data.imputed$LOV))[1]
data.imputed %>%
  select(WAITTIME.imputed) %>%
  filter(is.na(WAITTIME.imputed))
prop.table(table(data$WAITTIME))[2]
prop.table(table(data.imputed$WAITTIME.imputed))[1:2]

data.design.imputed <- svydesign(id = ~CPSUM, strata = ~CSTRATM, weight = ~PATWT, data = data.imputed, nest = TRUE)
(result.2a.imputed <- svymean(~WAITTIME.imputed, data.design.imputed, na.rm = T))
```


The average waiting time is almost 45 min. with a SE of 1.8512 --> interpretation
Show the proportion of missings --> MCAR? about 6.6. percent, and 2.7 percent not applicable (2.7 have probably too long or felt better) --> 6.6 percent can be imputed with proportion technique about 5 decreasesd show in presentation -_> comparison


#### b. Exclude the cluster and strata information, (redefine the svydesign), how this change your result in 2a, explain?

```{r}
# Defining the design object but without any design specifications
data.design.exc <- svydesign(id = ~id, weight = ~PATWT, data = data) 

(result.2b <- svymean(~WAITTIME, data.design.exc, na.rm = T))
```

The estimates are the same (about 45 min) but the the SE are too small for the option without any design specifications. Explanation needed.

#### c. Draw an SRS sample to estimate the average waiting time. Estimate the average waiting time, and compare the result(s) with the result(s) of 2a and 2b.


```{r}
# Drawing three srs samples from our original dataset (n = 1000, 5000, 10000) and calculating the final weights
srs.names <- c("SRS.1000", "SRS.5000", "SRS.10000")
srs.sizes <- c(1000, 5000, 10000)
srs.samples.vec <- pmap(list(x = length(srs.sizes), y = srs.sizes), ~srswr(n = .y, N = length(data$id)))

srs.samples.data <- map(srs.samples.vec, ~getdata(data, .x)) %>% 
  map2(., srs.sizes, ~mutate(.x, final.weights = PATWT * (34936 / .y)))

# Checking if the the new combined weights approximately add up to the original N 
map(srs.samples, ~sum(.$final.weights)) 

# Create for each SRS sample a svydesign object with the specific weights, excluding strata and cluster info
srs.samples.designs <- map(srs.samples.data, ~svydesign(id = ~.$id, weight = ~.$final.weights, data = .x, nest = TRUE)) %>% 
  set_names(nm = srs.names)

# Calculating the mean estimate of WAITTIME of the SRS samples and comparing them to the designs from before
results.srs <- map(srs.samples.designs, ~svymean(~WAITTIME, .x, na.rm = T))

# Comparing the results 
name.results <- c("2a Estimate (design specified)", "2b Estimate (design not specified)",
                  "ec Estimate SRS n = 1000", "2c Estimate SRS n = 5000", "2c Estimate SRS n = 10000")

# Still need to add the SEs to the table for better comparison
results.2 <- tibble(name.results, 
                    estimates = c(result.2a, result.2b, results.srs[[1]], results.srs[[2]], results.srs[[3]]),
                    SE = c(map_dbl(list(result.2a, result.2b, results.srs[[1]], results.srs[[2]], results.srs[[3]]), 
                                   ~sqrt(attr(., which = "var")))),
                    Bias = estimates - result.2a[[1]]) # add design effect

```


#### d. Draw a stratified sample to investigate the variability in waiting times over regions (Northeast, Midwest, South, West). Is there variability in waiting times over regions?


```{r}
# In order for drawing a stratified sample we need to arrange the original dataset in ascending strata order (Region)
data.strata.prep <- data %>%
  arrange(REGION)

# With the strata function we can draw a stratified sample. The size vector specifies the strata sample sizes 
# Implemented PPS allocation for now and n = 10000
strata.sizes <- data %>%
  group_by(REGION) %>%
  summarise(prop = (n() / 34936)) %>%
  mutate(strata.size = round(prop * 10000)) %>%
  mutate(population.size = prop * 34936) 

stratified.sample.vec <- strata(data = data.strata.prep, stratanames = "REGION", 
                                size = strata.sizes$strata.size, method = "srswr")
stratified.sample <- getdata(data.strata.prep, stratified.sample.vec) %>%
  left_join(strata.sizes, by = "REGION") %>% # join strata specific sizes (4 levels for 4 regions)
  mutate(weights.strata = population.size / strata.size) %>% # calculating strata specific weights (Nh / nh)
  mutate(final.weights = PATWT * weights.strata) # Combining the original weights with the strata specific weights 

# Checking if the the new combined weights approximately add up to the original N 
sum(stratified.sample$final.weights) 

# Defining the stratified svydesign() object with new weights and no original cluster information
stratified.sample.design <- svydesign(id = ~id, strata = ~REGION, weights = ~final.weights, data = stratified.sample, nest = TRUE) 

# Calculating the mean estimate of WAITTIME of the stratified sample
svymean(~WAITTIME, stratified.sample.design, na.rm = T)

# Testing the variability between regions concerning waiting time
svyby(~WAITTIME, ~REGION, stratified.sample.design, svymean, deff = T)
```

imagine we want to do this study again but only for estimation of waiting time 


### 3. Explain that a ratio estimator can be used to improve the estimation of the average waiting time. 

#### b. Describe potential variables for ratio estimation, and choose one to compute a ratio estimate of the average weighting time. 

we describe multiple ones but we will probably choose the ambulance one

#### c. Give a 95% confidence interval for your preferable estimate of the average waiting time, and explain the result.



### 4. Investigate differences in waiting times among ethnic groups by fitting a linear regression model. Compare estimated results of a stratified sampling design and an unequal probability sampling design.

```{r}
# Testing the relationship between ethnic groups and their waiting times with three sample designs
# --> Stratified sample with original weights * strata weights (sample.2d)
# --> Original design with n = N with original weights (data.design)
# --> SRS with keeping orginal weights (sample2.c.2)

# Dropping missings for plots


data.no.missings <- data %>%
  mutate(Design = "Original Design") %>%
  mutate(final.weights = PATWT) %>%
  filter_at(vars(matches("ETHUN|WAITTIME")), ~ . %!in% c(-7, -9 ))
sample.2c.2.no.missings <- sample.2c.2 %>%
  mutate(Design = "SRS") %>%
  mutate(final.weights = PATWT) %>%
  filter_at(vars(matches("ETHUN|WAITTIME")), ~ . %!in% c(-7, -9 ))
sample.2d.no.missings <- sample.2d %>%
  mutate(Design = "Stratified Sample") %>%
  mutate(final.weights = weights.final) %>%
  filter_at(vars(matches("ETHUN|WAITTIME")), ~ . %!in% c(-7, -9 ))

datasets.combined <- bind_rows(data.no.missings, sample.2c.2.no.missings, sample.2d.no.missings) %>%
  filter_at(vars(matches("ETHUN|WAITTIME")), ~ . %!in% c(-7, -9 )) %>%
  # filter(Design == "SRS") %>%
  mutate(Ethnicity = factor(ETHUN, levels = c(1, 2), labels = c("Hispanic or Latino", "Not Hispanic or Latino"))) %>%
  filter(id %in% c("101", "572", "103", "670"))
  

(reg.plot.Q4 <- ggplot() +
  # geom_jitter(data = datasets.combined,
  #             aes(x = Ethnicity, y = WAITTIME, color = Design, weight = final.weights, size = weights.final)) +
  geom_smooth(data = datasets.combined, 
              aes(x = Ethnicity, y = WAITTIME, group = Design, color = Design, weight = final.weights), method = "lm", se = FALSE) +
  labs(title = "Waiting Time over different Ethnicities and Designs", subtitle = "Question Nr. 4",
       x = "", y = "Waiting Time in Minutes") +
    theme_bw() +
    theme(text = element_text(size = 15), 
          axis.text.x = element_text(size = 13, face = "bold")))

ggsave(filename = "Regression Plot Q4.png", reg.plot.Q4, device = "png")

reg.data.original <- data.no.missings %>%
  filter(Design == "Original Design")
reg.data.original.design <- svydesign(id = ~CPSUM, strata = ~CSTRATM, weight = ~final.weights, data = reg.data.original, nest = TRUE)

reg.data.srs <- sample.2c.2.no.missings %>%
  filter(Design == "SRS")
reg.data.srs.design <- svydesign(id = ~CPSUM, strata = ~CSTRATM, weight = ~final.weights, data = reg.data.srs, nest = TRUE)
reg.data.srs.design <- svydesign(id = ~id, weight = ~final.weights, data = reg.data.srs, nest = TRUE)

reg.data.stratified <- sample.2d.no.missings %>%
  filter(Design == "Stratified Sample")
reg.data.stratified.design <- svydesign(id = ~CPSUM, strata = ~CSTRATM, weight = ~final.weights, data = reg.data.stratified, nest = TRUE)

model.Q4.original <- svyglm(formula = WAITTIME ~ as.factor(ETHUN), design = reg.data.original.design, data = reg.data.original)
model.Q4.original$R2 <- 1 - model.Q4.original$deviance/model.Q4.original$null.deviance

model.Q4.SRS <- svyglm(formula = WAITTIME ~ as.factor(ETHUN), family = gaussian(link = "identity"), design = reg.data.srs.design, 
                       data = reg.data.srs)
summary(model.Q4.SRS)
model.Q4.SRS$R2 <- 1 - model.Q4.SRS$deviance/model.Q4.SRS$null.deviance 

model.Q4.stratified <- svyglm(formula = WAITTIME ~ as.factor(ETHUN), design = reg.data.stratified.design, data = reg.data.stratified)
model.Q4.stratified$R2 <- 1 - model.Q4.stratified$deviance/model.Q4.stratified$null.deviance 


name.regression <- c("Original Design", "SRS", "Stratified Design")
R2.table <- tibble(name.regression, 
                   R2 = c(model.Q4.original$R2, model.Q4.SRS$R2, model.Q4.stratified$R2))

svyttest(WAITTIME ~ as.factor(ETHUN), design = reg.data.original.design)
svyttest(WAITTIME ~ as.factor(ETHUN), design = reg.data.srs.design)
svyttest(WAITTIME ~ as.factor(ETHUN), design = reg.data.stratified.design)
  

```


```{r}
# unweighted 
data.design.unweighted <- svydesign(id = ~id, data = data) 

# Best estimate from complete design
original <- svyby(~WAITTIME, by = ~ETHUN, svymean, design = data.design) %>% 
  mutate(type = "orginal")
exclude <- svyby(~WAITTIME, by = ~ETHUN, svymean, design = data.design.exc) %>% 
  mutate(type = "exclude")
srs <- svyby(~WAITTIME, by = ~ETHUN, svymean, design = sample.2c.des.2) %>% 
  mutate(type = "srs")
stratified <- svyby(~WAITTIME, by = ~ETHUN, svymean, design = sample.2d.des) %>% 
  mutate(type = "stratified")
unweighted <- svyby(~WAITTIME, by = ~ETHUN, svymean, design = data.design.unweighted) %>% 
  mutate(type = "unweighted")

results.weights <- bind_rows(original, exclude, srs, stratified, unweighted) %>% 
  arrange(ETHUN, type)

data %>% 
  group_by(ETHUN) %>% 
  summarise(sum(PATWT), sum(PATWT) / sum(data$PATWT), n(), n() / 34936)

g <- data %>% 
  filter(WAITTIME %!in%  c(-7, -9)) %>% 
  ggplot() +
  geom_point(aes(x = PATWT, y = WAITTIME)) +
  geom_smooth(aes(x = PATWT, y = WAITTIME))

data %>% 
  ggplot() +
  geom_histogram(aes(log(WAITTIME)))

data.new <- data %>% 
  filter(WAITTIME %!in% c(-7, -9)) %>% 
  mutate(WAIT.stand = (WAITTIME * mean(WAITTIME)) / sd(WAITTIME))

data.new %>% 
  ggplot() +
  geom_histogram(aes(WAIT.stand))

```


two models: 
are the individual weights the defined survey weights?
are the stratified weights in comparison to the individual weights related to ethnicity? What do we base the strata on? are these the same as beforehand or are these new strata that we ourselves need to define?
is it just comparing a model with the cluster and strata info like in the beginning and one model without any info


### 5. Assume unequal probability sampling, can you identify the important predictor variables of waiting times of patients visiting an ED? What can be concluded?


We dont know the severity of the patients, probably the most important predictor..


one or two issues. not everyone has to present, slides are nice answering questions 

focus more on the logic of the analysis and have a small (methodologically focused) storyline