---
title: "Survey Statistics | Project NHAMCS"
author: "Rianne, Piet, Zhenwei, Marco"
date: "Dec 16, 2019"
output:
  html_document:
    code_folding: hide
    theme: cosmo
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---


```{r}
library(tidyverse)
library(lubridate)
library(survey)
library(sampling)
library(kableExtra)
library(svglite)
'%!in%' <- function(x,y)!('%in%'(x,y))

set.seed(3)
```


```{r}
data <- read_csv("nhamcsed2010-short.csv") %>%
  rownames_to_column("id") %>% 
  mutate_at(vars(-LOV, -WAITTIME), ~replace(., . < 0, NA)) # Defining almost all "Missings" as NA only keeping them in LOV and WAITTIME

```


### 1. The variable CPSUM represents geographic primary sampling units (clusters), a variable CSTRATM represents relevant strata (e.g., emergency service areas within emergency departments or outpatient departments), variable PATWT represents an additional patient visit weight variable.

#### a. Explain this multistage sampling design. Do the PATWT weights sum up to N, and what does this mean?

```{r}
sum(data$PATWT)
```

Yes, the weights add up to N.

#### b. Define the svydesign() function with clusters, strata, and (additional) weights.

```{r}
data.design <- svydesign(id = ~CPSUM, strata = ~CSTRATM, weight = ~PATWT, data = data)
```

#### c. When considering gender differences in emergency visits, is it more likely for men than women to visit an emergency department, explain?

```{r}
# Population estimates for Males and Females 
svytotal(~interaction(SEX), data.design) 
```


#### d. Is there a mean difference between males and females in weights (PATWT), what do you conclude from this?

```{r}
# Males (2) and Females (1) proportion
data %>% 
  group_by(SEX) %>% 
  summarise(n = n(), mean.weight = mean(PATWT))
```


#### Further discussing the weights --> Piet can you make a nicer graph please

```{r}
weight.summary <- data %>% 
  group_by(PATWT) %>% 
  summarise(number = n()) %>% 
  arrange(PATWT)

n_distinct(weight.summary$PATWT)

data %>% 
  ggplot() +
  geom_histogram(aes(x = PATWT))
```


### 2. Estimate the average waiting time in the population under the specified svydesign. 

#### a. What do you conclude?

```{r}
(result.2a <- svymean(~WAITTIME, data.design, na.rm = T))
```

Imputing data by Zhenwei
```{r}
# Analysis of potentially imputable observations
imputation.table <- data %>% 
  mutate(helper.impute = case_when(WAITTIME == -9 & LOV == -9 ~ "Missing, but not imputable",
                                   WAITTIME == -9 & LOV != -9 ~ "Missing, but imputable",
                                   WAITTIME == -7 ~ "not applicable",
                                   TRUE ~ "Observed")) %>% 
  group_by(helper.impute) %>% 
  summarise(percentage.sample = round( (n()/34936) * 100, 2),
            percentage.population = round((sum(PATWT)/129843377) * 100, 2),
            sample.total = n(),
            population.total = sum(PATWT))

# Table for Appendix 
write.csv2(file = "results/imputatation.table.csv", x = imputation.table, row.names = F)

                                
# Imputing values single imputation by calculating the average waiting time proportion and multiplying it by the individual LOV time
data.imputed <- data %>%
  mutate(WAITTIME.prop = case_when(WAITTIME != -9 & WAITTIME != -7 & LOV != -9 ~ WAITTIME/LOV,
                                 TRUE ~ NA_real_)) %>%
  mutate(WAITTIME.imputed = case_when(WAITTIME != -9 & WAITTIME != -7 ~ WAITTIME,
                                      WAITTIME == -7 ~ -7,
                                      WAITTIME == -9 & LOV != -9 ~ LOV * weighted.mean(x = WAITTIME.prop, w = PATWT,
                                                                                       na.rm = T))) %>% 
  mutate(impute.group = case_when(WAITTIME != -9 & WAITTIME != -7 ~ "Observed, not imputed",
                                  WAITTIME == -7  ~ "Not applicable",
                                  WAITTIME == -9 & LOV != -9 ~ "Imputed",
                                  TRUE ~ "Missing, but not imputable")) %>% 
  mutate(WAITTIME.imputed = case_when(WAITTIME.imputed %in% c(-7, -9) ~ NA_real_,
                                      TRUE ~ WAITTIME.imputed))

# summary statistics for waiting time proportion
weighted.mean(data.imputed$WAITTIME.prop, w = data.imputed$PATWT, na.rm = T)

# new design with imputed values for waiting time
data.design.imputed <- svydesign(id = ~CPSUM, strata = ~CSTRATM, weight = ~PATWT, data = data.imputed)

data.imputed.filter <- data.imputed %>% 
  filter(impute.group == "Imputed")
data.design.imputed.filter <- svydesign(id = ~CPSUM, strata = ~CSTRATM, weight = ~PATWT, data = data.imputed.filter)

# comparing imputation strategy with original data 
(result.2a <- svymean(~WAITTIME, data.design, na.rm = T))
# Only the imputed ones
(result.2a.imputed.filter <- svymean(~WAITTIME.imputed, data.design.imputed.filter, na.rm = T))
# combined 
(result.2a.imputed <- svymean(~WAITTIME.imputed, data.design.imputed, na.rm = T))

# analysing the weights concerning the imputing.group variable
imputation.comparison.table <- data.imputed %>% 
  mutate(LOV.new = case_when(LOV == -9 ~ NA_real_,
                             TRUE ~ LOV)) %>% 
  group_by(impute.group) %>% 
  summarise(mean.weight = mean(PATWT),
            mean.WAITTIME = mean(WAITTIME.imputed, na.rm = T))
write.csv2(x = imputation.comparison.table, file = "results/imputation.comparison.table.csv")

```


The average waiting time is almost 45 min. with a SE of 1.8512 --> interpretation
Show the proportion of missings --> MCAR? about 6.6. percent, and 2.7 percent not applicable (2.7 have probably too long or felt better) --> 6.6 percent can be imputed with proportion technique about 5 decreasesd show in presentation -_> comparison


#### b. Exclude the cluster and strata information, (redefine the svydesign), how this change your result in 2a, explain?

```{r}
# Defining the design object but without any design specifications
data.design.exc <- svydesign(id = ~id, weight = ~PATWT, data = data.imputed) 

(result.2b <- svymean(~WAITTIME.imputed, data.design.exc, na.rm = T))
```

The estimates are the same (about 45 min) but the the SE are too small for the option without any design specifications. Explanation needed.

#### c. Draw an SRS sample to estimate the average waiting time. Estimate the average waiting time, and compare the result(s) with the result(s) of 2a and 2b.


```{r}
# Drawing three srs samples from our original dataset (n = 1000, 5000, 10000) and calculating the final weights
srs.names <- c("SRS.1000", "SRS.5000", "SRS.10000")
srs.sizes <- c(1000, 5000, 10000)
set.seed(347)
srs.samples.vec <- pmap(list(x = length(srs.sizes), y = srs.sizes), ~srswr(n = .y, N = length(data.imputed$id)))

srs.samples.data <- map(srs.samples.vec, ~getdata(data.imputed, .x)) %>% 
  map2(., srs.sizes, ~mutate(.x, final.weights = PATWT * (34936 / .y)))

# Checking if the the new combined weights approximately add up to the original N 
# map(srs.samples, ~sum(.$final.weights)) 

# Create for each SRS sample a svydesign object with the specific weights, excluding strata and cluster info
srs.samples.designs <- map(srs.samples.data, ~svydesign(id = ~.$id, weight = ~.$final.weights, data = .x, nest = TRUE)) %>% 
  set_names(nm = srs.names)

# Calculating the mean estimate of WAITTIME of the SRS samples and comparing them to the designs from before
results.srs <- map(srs.samples.designs, ~svymean(~WAITTIME.imputed, .x, na.rm = T))

# Comparing the results 
name.results <- c("2a Estimate (design specified)", "2b Estimate (design not specified)",
                  "2c Estimate SRS n = 1000", "2c Estimate SRS n = 5000", "2c Estimate SRS n = 10000")

# Still need to add the SEs to the table for better comparison
results.2 <- tibble(name.results, 
                    estimates = c(result.2a.imputed, result.2b, results.srs[[1]], results.srs[[2]], results.srs[[3]]),
                    SE = c(map_dbl(list(result.2a.imputed, result.2b, results.srs[[1]], results.srs[[2]], results.srs[[3]]), 
                                   ~sqrt(attr(., which = "var")))),
                    Bias = estimates - result.2a.imputed[[1]]) # add design effect

```


#### d. Draw a stratified sample to investigate the variability in waiting times over regions (Northeast, Midwest, South, West). Is there variability in waiting times over regions?


--> hier equal pps und optimal allocation () einfügen und zusammenfügen mit der tabelle davor lecture week stratified sampling 
```{r}
# In order for drawing a stratified sample we need to arrange the original dataset in ascending strata order (Region)
data.strata.prep <- data.imputed %>%
  arrange(REGION)

strata.sizes <- data.imputed %>%
  group_by(REGION) %>%
  summarise(prop = (n() / 34936)) %>%
  mutate(strata.size.pps = round(prop * 10000)) %>%
  mutate(population.size = prop * 34936) 

set.seed(87)
# Equal sizes
stratified.equal.sample.vec <- strata(data = data.strata.prep, stratanames = "REGION", 
                                      size = c(2500, 2500, 2500, 2500), method = "srswr")
stratified.equal.sample <- getdata(data.strata.prep, stratified.equal.sample.vec) %>%
  left_join(strata.sizes, by = "REGION") %>% # join strata specific sizes (4 levels for 4 regions)
  mutate(weights.strata = population.size / 2500) %>% # calculating strata specific weights (Nh / nh) -_> equal between strata
  mutate(final.weights = PATWT * weights.strata) # Combining the original weights with the strata specific weights
stratified.equalSize.sample.design <- svydesign(id = ~id, strata = ~REGION, weights = ~final.weights, data = stratified.equal.sample) 


#PPS
# With the strata function we can draw a stratified sample. The size vector specifies the strata sample sizes 
# Implemented PPS allocation for now and n = 10000

set.seed(87)
stratified.PPS.sample.vec <- strata(data = data.strata.prep, stratanames = "REGION", 
                                    size = strata.sizes$strata.size.pps, method = "srswr")
stratified.PPS.sample <- getdata(data.strata.prep, stratified.PPS.sample.vec) %>%
  left_join(strata.sizes, by = "REGION") %>% # join strata specific sizes (4 levels for 4 regions)
  mutate(weights.strata = population.size / strata.size.pps) %>% # calculating strata specific weights (Nh / nh)
  mutate(final.weights = PATWT * weights.strata) # Combining the original weights with the strata specific weights 
stratified.PPS.sample.design <- svydesign(id = ~id, strata = ~REGION, weights = ~final.weights, data = stratified.PPS.sample) 



# Neyman allocation
sd.neyman <- data.imputed %>% 
  group_by(REGION) %>% 
  summarise(sd.region = sd(WAITTIME.imputed, na.rm = T))

neyman.region.1 <- ( (8735 * sd.neyman$sd.region[1]) / 
                      ( (8735 * sd.neyman$sd.region[1]) + (7299 * sd.neyman$sd.region[2]) + (12458 * sd.neyman$sd.region[3]) + (6444 * sd.neyman$sd.region[4]) ) * 10000 )
neyman.region.2 <- ( (7299 * sd.neyman$sd.region[2]) / 
                       ( (8735 * sd.neyman$sd.region[1]) + (7299 * sd.neyman$sd.region[2]) + (12458 * sd.neyman$sd.region[3]) + (6444 * sd.neyman$sd.region[4]) ) * 10000 )
neyman.region.3 <- ( (12458 * sd.neyman$sd.region[3]) / 
                       ( (8735 * sd.neyman$sd.region[1]) + (7299 * sd.neyman$sd.region[2]) + (12458 * sd.neyman$sd.region[3]) + (6444 * sd.neyman$sd.region[4]) ) * 10000 )
neyman.region.4 <- ( (6444 * sd.neyman$sd.region[4]) / 
                       ( (8735 * sd.neyman$sd.region[1]) + (7299 * sd.neyman$sd.region[2]) + (12458 * sd.neyman$sd.region[3]) + (6444 * sd.neyman$sd.region[4]) ) * 10000 )
neyman.region.sizes <- c(neyman.region.1, neyman.region.2, neyman.region.3, neyman.region.4)
strata.sizes$strata.size.neyman <- round(neyman.region.sizes, 0)

set.seed(87)
stratified.neyman.sample.vec <- strata(data = data.strata.prep, stratanames = "REGION", 
                                       size = neyman.region.sizes, method = "srswr")
stratified.neyman.sample <- getdata(data.strata.prep, stratified.neyman.sample.vec) %>%
  left_join(strata.sizes, by = "REGION") %>% # join strata specific sizes (4 levels for 4 regions)
  mutate(weights.strata = population.size / strata.size.neyman) %>% # calculating strata specific weights (Nh / nh)
  mutate(final.weights = PATWT * weights.strata) # Combining the original weights with the strata specific weights
stratified.neyman.sample.design <- svydesign(id = ~id, strata = ~REGION, weights = ~final.weights, data = stratified.neyman.sample) 

stratfied.samples <- list(stratified.equalSize.sample.design, stratified.PPS.sample.design, stratified.neyman.sample.design)


# Calculating the mean estimate of WAITTIME of the SRS samples and comparing them to the designs from before
results.stratified <- map(stratfied.samples, ~svymean(~WAITTIME.imputed, .x, na.rm = T))

# Comparing the results 
name.stratified.results <- c("Equal sizes", "PPS", "Neyman")

# Still need to add the SEs to the table for better comparison
results.stratified.2 <- tibble(name.stratified.results, 
                               estimates = c(results.stratified[[1]], results.stratified[[2]], results.stratified[[3]]),
                               SE = map_dbl(results.stratified, 
                                              ~sqrt(attr(., which = "var"))),
                               Bias = estimates - result.2a.imputed[[1]]) # add design effect

##--> put into tabel from above

# Checking if the the new combined weights approximately add up to the original N 
sum(stratified.sample$final.weights) 

# Calculating the mean estimate of WAITTIME of the stratified sample
svymean(~WAITTIME.imputed, stratified.neyman.sample.design, na.rm = T)

# Testing the variability between regions concerning waiting time
regional.estimate <- svyby(~WAITTIME.imputed, ~REGION, stratified.neyman.sample.design, svymean, deff = T, na.rm = T) # check this 
write.csv2(x = regional.estimate, file = "results/regional.estimate.csv")
```

imagine we want to do this study again but only for estimation of waiting time 


### 3. Explain that a ratio estimator can be used to improve the estimation of the average waiting time. 

#### b. Describe potential variables for ratio estimation, and choose one to compute a ratio estimate of the average weighting time. 

```{r}
ratio.3 <- svyratio(~WAITTIME.imputed, ~ARREMS, design = data.design.imputed, na.rm = T)
ratio.3.predict <- predict(ratio.3, total = mean(data.imputed$ARREMS, na.rm = T), interval = "confidence")
```


ratio ambulance groups on waiting time

#### c. Give a 95% confidence interval for your preferable estimate of the average waiting time, and explain the result.

```{r}
ci.lb <- ratio.3.predict$total - (ratio.3.predict$se * 1.96) 
ci.ub <- ratio.3.predict$total + (ratio.3.predict$se * 1.96) 
ci <- c(ci.lb, ci.ub)
```



### 4. Investigate differences in waiting times among ethnic groups by fitting a linear regression model. Compare estimated results of a stratified sampling design and an unequal probability sampling design.

DV = ETHUN, IV = WAITTIME.imputed
--> Baseline model = data.design.imputed
--> Stratified sample = neyman allocation
--> unqueal probs = srs.samples.data[3]

--> Check conditions
- Homoscedasticity
- independence limitations
- normal residuals
- outlier control --> cooks distance, leverage

--> maybe plot and regression table

```{r}
# Testing the relationship between ethnic groups and their waiting times with three sample designs
# --> Stratified sample with original weights * strata weights (sample.2d)
# --> Original design with n = N with original weights (data.design)
# --> SRS with keeping orginal weights (sample2.c.2)

# Dropping missings for plots

## --> Baseline model Q4
mod.baseline.w.outlier <- data.imputed
mod.baseline.wo.outlier <- data.imputed %>% 
  filter(WAITTIME.imputed <= 600)

mod.baseline.w.outlier.design <- svydesign(id = ~CPSUM, strata = ~CSTRATM, weight = ~PATWT, data = mod.baseline.w.outlier)

mod.baseline.w.outlier.wo.design <- svydesign(id = ~CPSUM, strata = ~CSTRATM, weight = ~PATWT, data = mod.baseline.wo.outlier)

(model.Q4.original <- svyglm(formula = WAITTIME.imputed ~ as.factor(ETHUN), design = mod.baseline.w.outlier.design, 
                            data = mod.baseline.w.outlier))
(model.Q4.original$R2 <- 1 - model.Q4.original$deviance/model.Q4.original$null.deviance)

(model.Q4.original.wo.outlier <- svyglm(formula = WAITTIME.imputed ~ as.factor(ETHUN), design = mod.baseline.w.outlier.wo.design, 
                                       data = mod.baseline.wo.outlier))
(model.Q4.original.wo.outlier$R2 <- 1 - model.Q4.original.wo.outlier$deviance/model.Q4.original.wo.outlier$null.deviance)



## --> SRS model 10000

mod.srs <- srs.samples.designs[[3]]

(model.Q4.srs <- svyglm(formula = WAITTIME.imputed ~ as.factor(ETHUN), design = mod.srs, 
                            data = data.imputed))
(model.Q4.srs$R2 <- 1 - model.Q4.srs$deviance/model.Q4.srs$null.deviance)


## --> Neymann 10000

mod.neyman <- stratified.neyman.sample.design

(model.Q4.neyman <- svyglm(formula = WAITTIME.imputed ~ as.factor(ETHUN), design = mod.neyman, 
                            data = data.imputed))
(model.Q4.neyman$R2 <- 1 - model.Q4.neyman$deviance/model.Q4.neyman$null.deviance)


## plots

datasets.combined <- bind_rows(data.no.missings, sample.2c.2.no.missings, sample.2d.no.missings) %>%
  filter_at(vars(matches("ETHUN|WAITTIME")), ~ . %!in% c(-7, -9 )) %>%
  # filter(Design == "SRS") %>%
  mutate(Ethnicity = factor(ETHUN, levels = c(1, 2), labels = c("Hispanic or Latino", "Not Hispanic or Latino")))

(reg.plot.Q4 <- ggplot() +
  # geom_jitter(data = datasets.combined,
  #             aes(x = Ethnicity, y = WAITTIME, color = Design, weight = final.weights, size = weights.final)) +
  geom_smooth(data = datasets.combined, 
              aes(x = Ethnicity, y = WAITTIME, group = Design, color = Design, weight = final.weights), method = "lm", se = FALSE) +
  labs(title = "Waiting Time over different Ethnicities and Designs", subtitle = "Question Nr. 4",
       x = "", y = "Waiting Time in Minutes") +
    theme_bw() +
    theme(text = element_text(size = 15), 
          axis.text.x = element_text(size = 13, face = "bold")))
  

```


two models: 
are the individual weights the defined survey weights?
are the stratified weights in comparison to the individual weights related to ethnicity? What do we base the strata on? are these the same as beforehand or are these new strata that we ourselves need to define?
is it just comparing a model with the cluster and strata info like in the beginning and one model without any info


### 5. Assume unequal probability sampling, can you identify the important predictor variables of waiting times of patients visiting an ED? What can be concluded?


```{r}
data.Q5 %>% 
  ggplot() +
  geom_jitter(aes(x = RACEUN, y = WAITTIME.log)) 

a %>% 
  ggplot() +
  geom_histogram(aes(x = WAITTIME.imputed))

a %>% 
  group_by(RACEUN) %>% 
  summarise(mean(WAITTIME.imputed, na.rm = T))



data.Q5 <- data.imputed %>% 
  mutate(day.kind = case_when(VDAYR %in% c(2:6) ~ "Weekdays",
                              TRUE ~ "Weekend")) %>% 
  mutate(time = case_when(nchar(ARRTIME) == 4 ~ sub("(.{2})(.*)", "\\1:\\2", ARRTIME),
                          nchar(ARRTIME) == 3 ~ paste0("0", sub("(.{1})(.*)", "\\1:\\2", ARRTIME)),
                          nchar(ARRTIME) == 2 ~ paste0("00:", ARRTIME),
                          nchar(ARRTIME) == 1 ~ paste0("00:0", ARRTIME))) %>% 
  mutate(hour = substr(time, 1, 2)) %>% 
  mutate(time.kind = case_when(hour %in% c(7:23) ~ "Daytime",
                               TRUE ~ "Nighttime")) %>% 
  mutate(payment.combined = case_when(PAYPRIV == 1 ~ "Private",
                                      PAYMCAID == 1 | PAYMCARE == 1 & PAYPRIV == 0 ~ "Medicare/Medicaid",
                                      TRUE ~ "No expected source of Payment")) %>% 
  mutate(WAITTIME.log = log(WAITTIME.imputed + 1))


data.Q5.design <- svydesign(id = ~CPSUM, strata = ~CSTRATM, weight = ~PATWT, data = data.Q5)

(model.Q5.original <- svyglm(formula = WAITTIME.imputed ~ as.factor(payment.combined) + 
                               as.factor(time.kind) + as.factor(day.kind) +
                               as.factor(RESIDENCE) + as.factor(RACEUN),
                             design = data.Q5.design, 
                             data = data.Q5))
(model.Q5.original <- svyglm(formula = WAITTIME.log ~ as.factor(RACEUN),
                             design = data.Q5.design, 
                             data = data.Q5))
(model.Q5.original$R2 <- 1 - model.Q5.original$deviance/model.Q5.original$null.deviance)

hm(data.imputed$ARRTIME)
```


